---
title: "Reproducible documents"
author: "Kristina Hansen"
format: html
---

This is a test 

## About me 
- Name: Kristina 
- Age: 27
- Affiliation: PhD student 

I am doin a PhD about **preventing** type 2-diabetes among 30-50 years old adulit in *DK* an the *Faroe Islands*

## Simple code 
```{r simplecode}
3 * 3
```



```{r setup} 
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```

```{r test}
2 + 2
```



## Showing the data
```{r}
post_meal_data
glimpse(post_meal_data)
```

## Plot one continuous variable
```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Exercise: discrete plots
```{r}
#| fig-cap: "plot_Group"
#| label: fig-Group_bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-Group_bar above for a cool plot!

## Plotting two discrete variables
```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```
as.character fordi glykemi ikke er en kategorsik variabel som "fill" ellers kræver - dette løses med as.character

## plot uden at de er stacked ovenpå hianden (position_dodge)
```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

```{r}
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r}
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```
Fordi de to koder kommer i samme chunk placeres de ved sinden i hinanden ved "render"


## Plotting two continuous variables (+ geom_smooth for linjen)
```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose."
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Testing for Git
This section will be working of the Git-part of the workshop 
Opretter et Github repo
```{r}
# usethis::create_github_token()
```

Gemmmer kode 
```{r}
# gitcreds::gitcreds_set()
```
indtaster kode 


** her uploades hele projektet - måske jeg i nogle tilfølde ikke ønsker at uploade data (datasikkerhed og GDPR)

## Selecting data 
```{r}
select(post_meal_data, Age)

select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)
```

## Select helpers
```{r}
select(post_meal_data, starts_with("PG"))
```

```{r}
select(post_meal_data, ends_with("DXA"))
```
```{r}
# All columns containing the number "0"
select(post_meal_data, contains("0"))
```

## Renaming columns
```{r}
small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Training 
```{r}
# Task 1.
post_meal_data |>
  select(contains("OGGT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename()

# Task 4.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)
```

## Select and rename tasks
