---
title: "Reproducible documents"
author: "Kristina Hansen"
format: html
---

## About me 
- Name: Kristina 
- Age: 27
- Affiliation: PhD student 

I am doin a PhD about **preventing** type 2-diabetes among 30-50 years old adulit in *DK* an the *Faroe Islands*

## Simple code 
```{r simplecode}
3*3
```



```{r setup} 
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```

```{r test}
2 + 2
```


## Testing for Git
This section will be working of the Git-part of the workshop 

## Showing the data
```{r}
post_meal_data
glimpse(post_meal_data)
```

## Plot one continuous variable
```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Exercise: discrete plots
```{r}
#| fig-cap: "plot_Group"
#| label: fig-Group_bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-Group_bar above for a cool plot!

## Plotting two discrete variables
```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```
as.character fordi glykemi ikke er en kategorsik variabel som "fill" ellers kræver - dette løses med as.character

## plot uden at de er stacked ovenpå hianden (position_dodge)
```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```


